---
title: "Analysis for Manuscript: \nStage-mediated priority effects and species life history shape long-term competition dynamics"
author: "Heng-Xing Zou, Sebastian J. Schreiber, Volker H. W. Rudolf"
date: "10/20/2022"
output: html_document
---

# General Notes

- This file, when knitted, produces all figures included in the main text and Supporting Information. Note that the order might differ. 
- Before running, please make sure that this file is under the same folder with `Functions.R`.
- Some chunks may take very long time to run.
- Some chunks require changing certain places in `Functions.R`. Please follow instructions for those specific chunks.
- Please also read `IMPORTANT NOTES` at the beginning of `Function.R`.
- Please contact the author [Heng-Xing Zou](hz70@rice.edu) for questions or suggestions.

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = TRUE)
source("Functions.R")

fig_dir = "/Figures"
if (!dir.exists(fig_dir)) {
  dir.create(fig_dir)
}

```

# Data Generation

Data generation could take a long time. To jumpstart, run the following chunk to load the pre-saved environment.

```{r}

load("Aug242022.RData")

```

## Stage-mediated Interspecific Competition

- Invasion Analysis

Change `generate_amat()` in [Functions.R](Functions.R) accordingly: $\alpha_{11}$ and $\alpha_{22}$ should be constants (`"c"`), $\alpha_{12}$ and $\alpha_{21}$ should be variable (`"v"`).

```{r}

delta_s_lst = seq(-4, 4, 1)
init_pop = cbind(c(0, 0, 0, 0, 0, 0, 3), c(0, 0, 0, 0, 0, 0, 3))
para = cbind(c(12, 0.8, 0.8, 0.8, 0.8, 0.8, 0.8), c(12, 0.8, 0.8, 0.8, 0.8, 0.8, 0.8))
mu = c(0.2, 0.2)
s = 7
season_lens = 6*c(2, 4, 8, 16, 20)+1
n_seasons = 50

# Low intraspecific competition, symmetric

lowintra = invasion_multi(n_seasons, init_pop, para, c(B, 0.10, 0.10), delta_s_lst, var = 0, season_lens)

# Mid intraspecific competition, asymmetric

midintra = invasion_multi(n_seasons, init_pop, para, c(B, 0.10, 0.12), delta_s_lst, var = 0, season_lens)

# High intraspecific competition, symmetric

highintra = invasion_multi(n_seasons, init_pop, para, c(B, 0.12, 0.12), delta_s_lst, var = 0, season_lens)

```

- Calculating ND and RFD

```{r}

nd_rfd = midintra %>% select(inv_i, sensitivity, delta_s, season_len) %>% pivot_wider(names_from = inv_i, values_from = sensitivity) %>%
  mutate(ND = 1-sqrt(`1`*`2`), RFD = sqrt(`1`/`2`)) %>% 
  mutate(outcome = ifelse(1-ND < RFD & RFD < 1/(1-ND), "3", 
                          ifelse(RFD > 1-ND & RFD > 1/(1-ND), "2", 
                                 ifelse(RFD < 1-ND & RFD < 1/(1-ND), "1", "4"))))

```

- Periodic Environment

```{r}

init_pop = cbind(c(0, 0, 0, 0, 0, 0, 3), c(0, 0, 0, 0, 0, 0, 3))
para = cbind(c(12, 0.8, 0.8, 0.8, 0.8, 0.8, 0.8), c(12, 0.8, 0.8, 0.8, 0.8, 0.8, 0.8))
mu = c(0.2, 0.2)
s = 7
season_lens = 6*c(2, 4, 8, 16, 20)+1
n_seasons = 50

delta_s_lst = seq(-2, 2, 1)
var2 = invasion_multi(n_seasons, init_pop, para, c(B, 0.10, 0.12), delta_s_lst, var = 2, season_lens, periodic = T)

```

## Stage-mediated Interspecific Competition

Change `generate_amat()` in [Functions.R](Functions.R) accordingly: $\alpha_{11}$ and $\alpha_{22}$ should be constants (`"c"`), $\alpha_{12}$ and $\alpha_{21}$ should be variable (`"v"`).

- Invasion Analysis

```{r}

delta_s_lst = seq(-4, 4, 1)
init_pop = cbind(c(0, 0, 0, 0, 0, 0, 3), c(0, 0, 0, 0, 0, 0, 3))
para = cbind(c(12, 0.8, 0.8, 0.8, 0.8, 0.8, 0.8), c(12, 0.8, 0.8, 0.8, 0.8, 0.8, 0.8))
mu = c(0.2, 0.2)
s = 7
season_lens = 6*c(2, 4, 8, 16, 20)+1
n_seasons = 50

# Low intraspecific competition, symmetric

lowintra = invasion_multi(n_seasons, init_pop, para, c(B, 0.10, 0.10), delta_s_lst, var = 0, season_lens)

# Mid intraspecific competition, asymmetric

midintra = invasion_multi(n_seasons, init_pop, para, c(B, 0.10, 0.12), delta_s_lst, var = 0, season_lens)

# High intraspecific competition, symmetric

highintra = invasion_multi(n_seasons, init_pop, para, c(B, 0.12, 0.12), delta_s_lst, var = 0, season_lens)

```

- Calculating ND and RFD

```{r}

nd_rfd = midintra %>% select(inv_i, sensitivity, delta_s, season_len) %>% pivot_wider(names_from = inv_i, values_from = sensitivity) %>%
  mutate(ND = 1-sqrt(`1`*`2`), RFD = sqrt(`1`/`2`)) %>% 
  mutate(outcome = case_when(RFD > 1-ND & RFD < 1/(1-ND) ~ "3", 
                             RFD > 1-ND & RFD > 1/(1-ND) ~ "2", 
                             RFD < 1-ND & RFD < 1/(1-ND) ~ "1", 
                             RFD < 1-ND & RFD > 1/(1-ND) ~ "4"))

```

- Periodic Environment

```{r}

init_pop = cbind(c(0, 0, 0, 0, 0, 0, 3), c(0, 0, 0, 0, 0, 0, 3))
para = cbind(c(12, 0.8, 0.8, 0.8, 0.8, 0.8, 0.8), c(12, 0.8, 0.8, 0.8, 0.8, 0.8, 0.8))
mu = c(0.2, 0.2)
s = 7
season_lens = 6*c(2, 4, 8, 16, 20)+1
n_seasons = 50

delta_s_lst = seq(-2, 2, 1)
var2 = invasion_multi(n_seasons, init_pop, para, c(B, 0.10, 0.12), delta_s_lst, var = 2, season_lens, periodic = T)

```

## No Stage-mediated Interspecific Competition

Change `generate_amat()` in [Functions.R](Functions.R) accordingly: $\alpha_{11}$, $\alpha_{22}$, $\alpha_{12}$ and $\alpha_{21}$ should all be constants (`"c"`).

- Invasion Analysis

```{r}

delta_s_lst = seq(-4, 4, 1)
init_pop = cbind(c(0, 0, 0, 0, 0, 0, 3), c(0, 0, 0, 0, 0, 0, 3))
para = cbind(c(12, 0.8, 0.8, 0.8, 0.8, 0.8, 0.8), c(12, 0.8, 0.8, 0.8, 0.8, 0.8, 0.8))
mu = c(0.2, 0.2)
s = 7
season_lens = 6*c(2, 4, 8, 16, 20)+1
n_seasons = 50

# Low intraspecific competition, symmetric

xlowintra = invasion_multi(n_seasons, init_pop, para, c(B, 0.10, 0.10), delta_s_lst, var = 0, season_lens)

# Mid intraspecific competition, asymmetric

xmidintra = invasion_multi(n_seasons, init_pop, para, c(B, 0.10, 0.12), delta_s_lst, var = 0, season_lens)

# High intraspecific competition, symmetric

xhighintra = invasion_multi(n_seasons, init_pop, para, c(B, 0.12, 0.12), delta_s_lst, var = 0, season_lens)

```

- Calculating ND and RFD

```{r}

xnd_rfd = xmidintra %>% select(inv_i, sensitivity, delta_s, season_len) %>% pivot_wider(names_from = inv_i, values_from = sensitivity) %>%
  mutate(ND = 1-sqrt(`1`*`2`), RFD = sqrt(`1`/`2`)) %>% 
  mutate(outcome = case_when(RFD > 1-ND & RFD < 1/(1-ND) ~ "3", 
                             RFD > 1-ND & RFD > 1/(1-ND) ~ "2", 
                             RFD < 1-ND & RFD < 1/(1-ND) ~ "1", 
                             RFD < 1-ND & RFD > 1/(1-ND) ~ "4"))

```

- Periodic Environment

```{r}

init_pop = cbind(c(0, 0, 0, 0, 0, 0, 3), c(0, 0, 0, 0, 0, 0, 3))
para = cbind(c(12, 0.8, 0.8, 0.8, 0.8, 0.8, 0.8), c(12, 0.8, 0.8, 0.8, 0.8, 0.8, 0.8))
mu = c(0.2, 0.2)
s = 7
season_lens = 6*c(2, 4, 8, 16, 20)+1
n_seasons = 50

delta_s_lst = seq(-2, 2, 1)
xvar2 = invasion_multi(n_seasons, init_pop, para, c(B, 0.10, 0.12), delta_s_lst, var = 2, season_lens, periodic = T)

```

# Main Text Figures

If mathematical expressions are not showing correctly, go to Tools -> Global Options -> Graphics, and select backend graphic device as "AGG".

## Figure 2

```{r}

compData = tibble(a12 = comp_coeff(B, scal, xmid, timeint, "v"), a21 = comp_coeff(B, -scal, xmid, timeint, "v"), pshift = timeint)

plcomp_inter =
compData %>% pivot_longer(col = a12:a21, names_to = "coefficient", values_to = "values") %>% 
  ggplot(aes(x = pshift, y = values, color = coefficient)) + geom_line() + geom_point(size = 2) + theme_bw() + ggtitle("A") + 
  scale_color_manual(name = "Coefficients", values = color_scheme, labels = c(expression(paste(beta[12])), expression(paste(beta[21])))) +
  scale_x_continuous(name = expression(paste("Initial stage difference (", Delta, "s)")), breaks = timeint) +
  ylab("Competition coefficients") +
  theme(panel.grid = element_blank()) + theme(axis.text = element_text(size = 15), axis.title = element_text(size = 20), 
                                              legend.text = element_text(size = 17), legend.title = element_text(size = 20), 
                                              plot.title = element_text(size = 20))

time = 130
season_len = 13

out = seasons(time, init_pop, para, alpha, mean_diff, var = 0, season_len)

pldyn = 
convert_output(out) %>% pivot_longer(col = 1:7, names_to = "stage", values_to = "population") %>% 
  group_by(time, species) %>% summarize(sum(population)) %>% rename(population = `sum(population)`) %>% 
  ggplot(aes(x = time, y = population, color = species)) + geom_line() + theme_bw() + 
  scale_color_manual(values = color_scheme) + ggtitle("B") + 
  scale_x_continuous(name = "Season", breaks = seq(0, time+1, season_len), label = seq(0, time/season_len, 1)) +
  geom_vline(xintercept = seq(season_len+1, time+1, season_len), color = "#577C8A", linetype = 4) +
  guides(color = guide_legend(title = "Species")) + ylab("Population") + xlab("Time") +
  theme(panel.grid.major.x = element_blank(), panel.grid.minor.x = element_blank()) + 
  theme(axis.text = element_text(size = 15), axis.title = element_text(size = 20), axis.ticks = element_blank(), 
        legend.text = element_text(size = 17), legend.title = element_text(size = 20), plot.title = element_text(size = 20))

```

- Save Figure 2

```{r}

ggsave("Figures/Fig2.pdf", plot = (plcomp_inter | pldyn), 
       width = 3600, height = 1200, unit = "px")

```

## Figure 3

- Deterministic (panels A, C)

```{r}

delta_s_lst = seq(-4, 4, 1)

plmid_tra = 
midintra %>% select(inv_i, lambda_ij, delta_s, season_len) %>% 
  pivot_wider(names_from = inv_i, values_from = lambda_ij) %>% 
  mutate(outcome = ifelse(`1`>0 & `2`>0, 3, 
                          ifelse(`1`>0 & `2`<0, 1, 
                                 ifelse(`1`<0 & `2`>0, 2, 4)))) %>% 
  ggplot(aes(x = as.factor(delta_s), y = as.factor(season_len), fill = factor(outcome))) + geom_tile(color = "white") + theme_bw() + 
  scale_x_discrete(name = expression(paste(Delta, "s")), breaks = delta_s_lst, label = delta_s_lst) +
  scale_y_discrete(name = "# Generations per season", breaks = season_lens, label = (season_lens-1)/6) +
  geom_rect(aes(xmin = 2.5, ymin = 0.5, xmax = 7.5, ymax = 5.5), color = "black", alpha = 0, size = 1) + 
  # change here; for periodic, guide goes here
  scale_fill_manual(name = "Outcome", labels = c("1 wins", "2 wins", "Coexistence", "Numeric PE"), values = color_scheme) +
  # scale_fill_manual(name = "Outcome", labels = c("1 wins", "2 wins", "Numeric PE"), values = c("#a61b29", "#0eb0c9", "#577C8A"), guide = "none") +
  theme(panel.grid.major = element_blank()) + 
  ggtitle(expression(paste("A. Constant ", Delta, "s"))) +
  theme(axis.text = element_text(size = 15), axis.ticks = element_blank(), 
        axis.title.x = element_blank(), axis.title.y = element_text(size = 20), 
        legend.text = element_text(size = 17), legend.title = element_text(size = 20), 
        plot.title = element_text(size = 20))

xplmid_tra = 
xmidintra %>% select(inv_i, lambda_ij, delta_s, season_len) %>% 
  pivot_wider(names_from = inv_i, values_from = lambda_ij) %>% 
  mutate(outcome = ifelse(`1`>0 & `2`>0, 3, 
                          ifelse(`1`>0 & `2`<0, 1, 
                                 ifelse(`1`<0 & `2`>0, 2, 4)))) %>% 
  ggplot(aes(x = as.factor(delta_s), y = as.factor(season_len), fill = factor(outcome))) + geom_tile(color = "white") + theme_bw() + 
  scale_x_discrete(name = expression(paste(Delta, "s")), breaks = delta_s_lst, label = delta_s_lst) +
  scale_y_discrete(name = "# Generations per season", breaks = season_lens, label = (season_lens-1)/6) +
  geom_rect(aes(xmin = 2.5, ymin = 0.5, xmax = 7.5, ymax = 5.5), color = "black", alpha = 0, size = 1) + 
  # change here 
  # scale_fill_manual(name = "Outcome", labels = c("1 wins", "2 wins", "Coexistence", "Numeric PE"), values = color_scheme, guide = "none") +
  scale_fill_manual(name = "Outcome", labels = c("1 wins", "2 wins", "Numeric PE"), values = c("#a61b29", "#0eb0c9", "#577C8A"), guide = "none") +
  theme(panel.grid.major = element_blank()) + 
  ggtitle(expression(paste("C. Constant ", Delta, "s"))) +
  theme(axis.text = element_text(size = 15), axis.ticks = element_blank(), 
        axis.title.x = element_text(size = 20), axis.title.y = element_text(size = 20), 
        legend.text = element_text(size = 17), legend.title = element_text(size = 20), plot.title = element_text(size = 20))

```

- Periodic (Panels B, D)

```{r}

delta_s_lst = seq(-2, 2, 1)

plmid_tra_peri = 
  var2 %>% select(inv_i, lambda_ij, delta_s, season_len) %>% 
  pivot_wider(names_from = inv_i, values_from = lambda_ij) %>% 
  mutate(outcome = ifelse(`1`>0 & `2`>0, 3, 
                          ifelse(`1`>0 & `2`<0, 1, 
                                 ifelse(`1`<0 & `2`>0, 2, 4)))) %>% 
  ggplot(aes(x = as.factor(delta_s), y = as.factor(season_len), fill = factor(outcome))) + geom_tile(color = "white") + theme_bw() + 
  scale_x_discrete(name = expression(bar(paste(Delta, "s"))), breaks = delta_s_lst, label = delta_s_lst) +
  scale_y_discrete(name = "# Generations per season", breaks = season_lens, label = (season_lens-1)/6) +
  # change here 
  scale_fill_manual(name = "Outcome", labels = c("1 wins", "2 wins", "Coexistence", "Numeric PE"), values = color_scheme, guide = "none") +
  # scale_fill_manual(name = "Outcome", labels = c("1 wins", "2 wins", "Numeric PE"), values = c("#a61b29", "#0eb0c9", "#577C8A"), guide = "none") +
  theme(panel.grid.major = element_blank()) + 
  ggtitle(expression(paste("B. Periodic ", Delta, "s"))) +
  theme(axis.text = element_text(size = 15), axis.title = element_blank(), axis.ticks = element_blank(), 
        legend.text = element_text(size = 17), legend.title = element_text(size = 20), plot.title = element_text(size = 20))

xplmid_tra_peri = 
  xvar2 %>% select(inv_i, lambda_ij, delta_s, season_len) %>% 
  pivot_wider(names_from = inv_i, values_from = lambda_ij) %>% 
  mutate(outcome = ifelse(`1`>0 & `2`>0, 3, 
                          ifelse(`1`>0 & `2`<0, 1, 
                                 ifelse(`1`<0 & `2`>0, 2, 4)))) %>% 
  ggplot(aes(x = as.factor(delta_s), y = as.factor(season_len), fill = factor(outcome))) + geom_tile(color = "white") + theme_bw() + 
  scale_x_discrete(name = expression(bar(paste(Delta, "s"))), breaks = delta_s_lst, label = delta_s_lst) +
  scale_y_discrete(name = "# Generations per season", breaks = season_lens, label = (season_lens-1)/6) +
  # change here 
  # scale_fill_manual(name = "Outcome", labels = c("1 wins", "2 wins", "Coexistence", "Numeric PE"), values = color_scheme, guide = "none") +
  scale_fill_manual(name = "Outcome", labels = c("1 wins", "Numeric PE"), values = c("#a61b29", "#577C8A"), guide = "none") +
  theme(panel.grid.major = element_blank()) + 
  ggtitle(expression(paste("D. Periodic ", Delta, "s"))) +
  theme(axis.text = element_text(size = 15), axis.title.x = element_text(size = 20), 
        axis.title.y = element_blank(), axis.ticks = element_blank(), 
        legend.text = element_text(size = 17), legend.title = element_text(size = 20), plot.title = element_text(size = 20))

```

- Save Figure 3

```{r}

ggsave("Figures/Fig3.pdf", 
       plot = plmid_tra + plmid_tra_peri + xplmid_tra + xplmid_tra_peri + 
              plot_layout(widths = c(1.75, 1), guides = "collect"), 
       width = 3000, height = 2250, unit = "px")

```

## Figure 4

- Stage-mediated Interspecific Competition

```{r}

ND_min = -0.3
ND_max = 0.6
RFD_min = -1.3
RFD_max = 1.3

# This is the only panel with legend

midintra_tra_13 =
  generate_baseplot(ND_min, ND_max, RFD_min, RFD_max) + 
  geom_path(data = nd_rfd %>% filter(season_len == 13), 
            aes(x = ND, y = log10(RFD))) + 
  geom_point(data = nd_rfd %>% filter(season_len == 13), 
             aes(x = ND, y = log10(RFD), color = outcome), size = 3) + 
  scale_color_manual(name = "Outcome", labels = c("1 wins", "2 wins", "Coexistence", "Numeric PE"), 
                     values = color_scheme) + 
  ggtitle("A. T = 2 generations") +
  annotate("text", size = 6,
           x = c(0.51, 0.51, -0.2, 0.4, 0.4), 
           y = c(0.12, -0.12, -0.2, 0.45, -0.45), 
           label = c("4", "-4", "0", "2", "-2")) + 
  theme(axis.text = element_text(size = 15), 
        axis.title.x = element_blank(), axis.title.y = element_text(size = 20),
        axis.ticks = element_blank(), plot.title = element_text(size = 20), 
        legend.text = element_text(size = 17), legend.title = element_text(size = 20))

  
ND_min = -0.2
ND_max = 0.25
RFD_min = -0.25
RFD_max = 0.25

midintra_tra_25 = 
  generate_baseplot(ND_min, ND_max, RFD_min, RFD_max) + 
  geom_path(data = nd_rfd %>% filter(season_len == 25), 
            aes(x = ND, y = log10(RFD))) + 
  geom_point(data = nd_rfd %>% filter(season_len == 25), 
             aes(x = ND, y = log10(RFD), color = outcome), size = 3) + 
  scale_color_manual(name = "Outcome", labels = c("1 wins", "2 wins", "Coexistence", "Numeric PE"), 
                     values = color_scheme, guide = "none") + 
  ggtitle("B. T = 4 generations") +
  annotate("text", size = 6,
           x = c(0.23, 0.23, -0.14, 0.12, 0.12),
           y = c(0.08, -0.14, -0.03, 0.08, -0.12),
           label = c("4", "-4", "0", "2", "-2")) + 
  theme(axis.text = element_text(size = 15), 
        axis.title.x = element_blank(), axis.title.y = element_blank(), 
        axis.ticks = element_blank(), plot.title = element_text(size = 20), 
        legend.text = element_text(size = 17), legend.title = element_text(size = 20))

```

- No Stage-mediated Interspecific Competition

```{r}

ND_min = -1
ND_max = 0.1
RFD_min = -1
RFD_max = 1

xmidintra_tra_13 =
  generate_baseplot(ND_min, ND_max, RFD_min, RFD_max) + 
  geom_path(data = xnd_rfd %>% filter(season_len == 13), 
            aes(x = ND, y = log10(RFD))) + 
  geom_point(data = xnd_rfd %>% filter(season_len == 13), 
             aes(x = ND, y = log10(RFD), color = outcome), size = 3) + 
  scale_color_manual(name = "Outcome", labels = c("1 wins", "2 wins", "Numeric PE"), 
                     values = c("#a61b29", "#0eb0c9", "#577C8A"), guide = "none") +
  ggtitle("C. T = 2 generations") +
  annotate("text", size = 6,
           x = c(-0.43, -0.41, -0.48, -0.62, -0.62), 
           y = c(0.4, -0.4, -0.1, 0.5, -0.55), 
           label = c("4", "-4", "0", "2", "-2")) + 
  theme(axis.text = element_text(size = 15), 
        axis.title.x = element_text(size = 20), axis.title.y = element_text(size = 20),
        axis.ticks = element_blank(), plot.title = element_text(size = 20), 
        legend.text = element_text(size = 17), legend.title = element_text(size = 20))

  
ND_min = -0.55
ND_max = 0.1
RFD_min = -0.2
RFD_max = 0.2

xmidintra_tra_25 =
  generate_baseplot(ND_min, ND_max, RFD_min, RFD_max) + 
  geom_path(data = xnd_rfd %>% filter(season_len == 25), 
            aes(x = ND, y = log10(RFD))) + 
  geom_point(data = xnd_rfd %>% filter(season_len == 25), 
             aes(x = ND, y = log10(RFD), color = outcome), size = 3) + 
  scale_color_manual(name = "Outcome", labels = c("Numeric PE"), 
                     values = c("#577C8A"), guide = "none") +
  ggtitle("D. T = 4 generations") + 
  annotate("text", size = 6,
           x = c(-0.41, -0.4, -0.5, -0.38, -0.37),
           y = c(0.09, -0.12, -0.02, 0, -0.03),
           label = c("4", "-4", "0", "2", "-2")) +
  theme(axis.text = element_text(size = 15), 
        axis.title.x = element_text(size = 20), axis.title.y = element_blank(),
        axis.ticks = element_blank(), plot.title = element_text(size = 20), 
        legend.text = element_text(size = 17), legend.title = element_text(size = 20))

```

- Save Figure 4

```{r}

ggsave("Figures/Fig4.pdf", 
       plot = (midintra_tra_13 | midintra_tra_25) / (xmidintra_tra_13 | xmidintra_tra_25) + 
         plot_layout(guides = "collect"), 
       width = 3000, height = 2250, unit = "px")

```

## Figure 5

- Stage-mediated Interspecific Competition

```{r}

vnd_rfd = var2 %>% 
  select(inv_i, sensitivity, delta_s, season_len) %>% 
  pivot_wider(names_from = inv_i, values_from = sensitivity) %>%
  mutate(ND = 1-sqrt(`1`*`2`), RFD = sqrt(`1`/`2`)) %>% 
  mutate(outcome = ifelse(1-ND < RFD & RFD < 1/(1-ND), "3", 
                          ifelse(RFD > 1-ND & RFD > 1/(1-ND), "2", 
                                 ifelse(RFD < 1-ND & RFD < 1/(1-ND), "1", "4"))))

# 2 Generations

ND_min = -0.3
ND_max = 0.6
RFD_min = -1.3
RFD_max = 1.3

midintra_tra_peri_13 =
  generate_baseplot(ND_min, ND_max, RFD_min, RFD_max) + 
  geom_path(data = vnd_rfd %>% filter(season_len == 13), 
            aes(x = ND, y = log10(RFD))) + 
  geom_point(data = vnd_rfd %>% filter(season_len == 13), 
             aes(x = ND, y = log10(RFD), color = outcome), size = 3) + 
  scale_color_manual(name = "Outcome", labels = c("1 wins", "2 wins", "Coexistence", "Numeric PE"), 
                     values = color_scheme) + 
  ggtitle("A. T = 2 generations") + 
  annotate("text", size = 6,
           x = c(0.55, 0.08, 0.08),
           y = c(-0.1, 0.24, -0.27),
           label = c("0", "2", "-2")) +
  theme(axis.text = element_text(size = 15), 
        axis.title.x = element_blank(), axis.title.y = element_text(size = 20),
        axis.ticks = element_blank(), plot.title = element_text(size = 20), 
        legend.text = element_text(size = 17), legend.title = element_text(size = 20))


# 4 Generations

ND_min = -0.2
ND_max = 0.25
RFD_min = -0.25
RFD_max = 0.25

midintra_tra_peri_25 =
  generate_baseplot(ND_min, ND_max, RFD_min, RFD_max) + 
  geom_path(data = vnd_rfd %>% filter(season_len == 25), 
            aes(x = ND, y = log10(RFD))) + 
  geom_point(data = vnd_rfd %>% filter(season_len == 25), 
             aes(x = ND, y = log10(RFD), color = outcome), size = 3) + 
  scale_color_manual(name = "Outcome", labels = c("1 wins", "2 wins", "Coexistence", "Numeric PE"), 
                     values = color_scheme, guide = "none") + 
  ggtitle("B. T = 4 generations") +
  annotate("text", size = 6,
           x = c(0.13, 0.03, 0.03),
           y = c(-0.02, 0.05, -0.10),
           label = c("0", "2", "-2")) +
  theme(axis.text = element_text(size = 15), 
        axis.title.x = element_blank(), axis.title.y = element_blank(),
        axis.ticks = element_blank(), plot.title = element_text(size = 20), 
        legend.text = element_text(size = 17), legend.title = element_text(size = 20))

```

- No Stage-mediated Interspecific Competition

```{r}

xvnd_rfd = xvar2 %>% 
  select(inv_i, sensitivity, delta_s, season_len) %>% 
  pivot_wider(names_from = inv_i, values_from = sensitivity) %>%
  mutate(ND = 1-sqrt(`1`*`2`), RFD = sqrt(`1`/`2`)) %>% 
  mutate(outcome = ifelse(1-ND < RFD & RFD < 1/(1-ND), "3", 
                          ifelse(RFD > 1-ND & RFD > 1/(1-ND), "2", 
                                 ifelse(RFD < 1-ND & RFD < 1/(1-ND), "1", "4"))))

ND_min = -1
ND_max = 0.1
RFD_min = -1
RFD_max = 1

xmidintra_tra_peri_13 =
  generate_baseplot(ND_min, ND_max, RFD_min, RFD_max) + 
  geom_path(data = xvnd_rfd %>% filter(season_len == 13), 
            aes(x = ND, y = log10(RFD))) + 
  geom_point(data = xvnd_rfd %>% filter(season_len == 13), 
             aes(x = ND, y = log10(RFD), color = outcome), size = 3) + 
  scale_color_manual(name = "Outcome", labels = c("1 wins", "Numeric PE"), 
                     values = c("#a61b29", "#577C8A"), guide = "none") +
  ggtitle("C. T = 2 generations") +
  annotate("text", size = 6,
           x = c(-0.41, -0.42, -0.43),
           y = c(-0.01, 0.25, -0.28),
           label = c("0", "2", "-2")) +
  theme(axis.text = element_text(size = 15), 
        axis.title.y = element_text(size = 20), axis.title.x = element_text(size = 20),
        axis.ticks = element_blank(), plot.title = element_text(size = 20), 
        legend.text = element_text(size = 17), legend.title = element_text(size = 20))

  
ND_min = -0.55
ND_max = 0.1
RFD_min = -0.2
RFD_max = 0.2

xmidintra_tra_peri_25 =
  generate_baseplot(ND_min, ND_max, RFD_min, RFD_max) + 
  geom_path(data = xvnd_rfd %>% filter(season_len == 25), 
            aes(x = ND, y = log10(RFD))) + 
  geom_point(data = xvnd_rfd %>% filter(season_len == 25), 
             aes(x = ND, y = log10(RFD), color = outcome), size = 3) + 
  scale_color_manual(name = "Outcome", labels = c("Numeric PE"), 
                     values = c("#577C8A"), guide = "none") +
  ggtitle("D. T = 4 generations") +
  annotate("text", size = 6,
           x = c(-0.38, -0.46, -0.46),
           y = c(-0.01, 0.04, -0.07),
           label = c("0", "2", "-2")) +
  theme(axis.text = element_text(size = 15), 
        axis.title.x = element_text(size = 20), axis.title.y = element_blank(),
        axis.ticks = element_blank(), plot.title = element_text(size = 20), 
        legend.text = element_text(size = 17), legend.title = element_text(size = 20))

```

- Save Figure 5

```{r}

ggsave("Figures/Fig5.pdf", 
       plot = (midintra_tra_peri_13 | midintra_tra_peri_25) / (xmidintra_tra_peri_13 | xmidintra_tra_peri_25) + 
         plot_layout(guides = "collect"), 
       width = 3000, height = 2250, unit = "px")

```

# Appendix II

```{r}

pllow_tra =
lowintra %>% select(inv_i, lambda_ij, delta_s, season_len) %>% 
  pivot_wider(names_from = inv_i, values_from = lambda_ij) %>% 
  mutate(outcome = ifelse(`1`>0 & `2`>0, 3, 
                          ifelse(`1`>0 & `2`<0, 1, 
                                 ifelse(`1`<0 & `2`>0, 2, 4)))) %>% 
  ggplot(aes(x = as.factor(delta_s), y = as.factor(season_len), fill = factor(outcome))) + geom_tile(color = "white") + theme_bw() + 
  scale_x_discrete(name = "Initial stage difference", breaks = delta_s_lst, label = delta_s_lst) +
  scale_y_discrete(name = "# Generations per season", breaks = season_lens, label = (season_lens-1)/6) +
  # change here 
  scale_fill_manual(name = "Outcome", labels = c("1 wins", "2 wins", "Coexistence", "Numeric PE"), values = color_scheme) +
  # scale_fill_manual(name = "Outcome", labels = c("1 wins", "2 wins", "Numeric PE"), values = c("#a61b29", "#0eb0c9", "#577C8A"), guide = "none") +
  theme(panel.grid.major = element_blank()) + ggtitle("A") +
  theme(axis.text = element_text(size = 15), axis.ticks = element_blank(), 
        axis.title.x = element_blank(), axis.title.y = element_text(size = 20), 
        legend.text = element_text(size = 17), legend.title = element_text(size = 20), plot.title = element_text(size = 20))

plhigh_tra =
highintra %>% select(inv_i, lambda_ij, delta_s, season_len) %>% 
  pivot_wider(names_from = inv_i, values_from = lambda_ij) %>% 
  mutate(outcome = ifelse(`1`>0 & `2`>0, 3, 
                          ifelse(`1`>0 & `2`<0, 1, 
                                 ifelse(`1`<0 & `2`>0, 2, 4)))) %>% 
  ggplot(aes(x = as.factor(delta_s), y = as.factor(season_len), fill = factor(outcome))) + geom_tile(color = "white") + theme_bw() + 
  scale_x_discrete(name = "Initial stage difference", breaks = delta_s_lst, label = delta_s_lst) +
  scale_y_discrete(name = "# Generations per season", breaks = season_lens, label = (season_lens-1)/6) +
  # change here 
  scale_fill_manual(name = "Outcome", labels = c("1 wins", "2 wins", "Coexistence", "Numeric PE"), values = color_scheme, guide = "none") +
  # scale_fill_manual(name = "Outcome", labels = c("1 wins", "2 wins", "Numeric PE"), values = c("#a61b29", "#0eb0c9", "#577C8A"), guide = "none") +
  theme(panel.grid.major = element_blank()) + ggtitle("B") +
  theme(axis.text = element_text(size = 15), axis.ticks = element_blank(), 
        axis.title.x = element_text(size = 20), axis.title.y = element_text(size = 20), 
        legend.text = element_text(size = 17), legend.title = element_text(size = 20), plot.title = element_text(size = 20))

```

## Figure S1

```{r}

ggsave("Figures/FigS1.png", plot = pllow_tra / plhigh_tra + plot_layout(guides = "collect"), 
       width = 2250, height = 2250, unit = "px")

```

# Appendix III

Change `generate_amat()` in [Functions.R](Functions.R) accordingly: $\alpha_{11}$, $\alpha_{22}$, $\alpha_{12}$ and $\alpha_{21}$ should all be variable (`"v"`).

## Competition-phenology Function

```{r}

timeint = seq(-4, 4, 1)
alpha = c(0.225, 0.10, 0.12)
scal = 0.85
xmid = 0

compData = tibble(a12 = comp_coeff(alpha[1], scal, xmid, timeint, "v"), a21 = comp_coeff(alpha[1], -scal, xmid, timeint, "v"), 
                  a11 = comp_coeff(alpha[2], scal, xmid, timeint, "v"), 
                  a22 = comp_coeff(alpha[3], scal, xmid, timeint, "v"), pshift = timeint)

plcomp_both =
compData %>% pivot_longer(col = a12:a22, names_to = "coefficient", values_to = "values") %>% 
  ggplot(aes(x = pshift, y = values, color = coefficient)) + 
  geom_line(aes(linetype = coefficient)) + geom_point(aes(shape = coefficient), size = 2) + theme_bw() + 
  scale_color_manual(name = "Coefficients", values = c("#c66d76", "#6a121b", "#097180", "#65ccdc"), 
                     labels = c(expression(paste(alpha[11])), expression(paste(beta[12])), 
                                expression(paste(beta[21])), expression(paste(alpha[22])))) +
  scale_shape_manual(name = "Coefficients", values = c(17, 16, 16, 17), 
                     labels = c(expression(paste(alpha[11])), expression(paste(beta[12])), 
                                expression(paste(beta[21])), expression(paste(alpha[22])))) + 
  scale_linetype_manual(name = "Coefficients", values = c(8, 1, 1, 8), 
                        labels = c(expression(paste(alpha[11])), expression(paste(beta[12])), 
                                expression(paste(beta[21])), expression(paste(alpha[22])))) +
  ylab("") + scale_x_continuous(name = "Initial stage difference", breaks = timeint) + ylab("Competition coefficients") +
  theme(panel.grid = element_blank()) + theme(axis.text = element_text(size = 15), axis.title = element_text(size = 20), 
                                                    legend.text = element_text(size = 17), legend.title = element_text(size = 20))

```

## Figure S2

```{r}

ggsave("Figures/FigS2.png", plot = plcomp_both, 
       width = 2000, height = 1200, unit = "px")

```

## Invasion Analysis

- Data Generation

```{r}

delta_s_lst = seq(-4, 4, 1)
init_pop = cbind(c(0, 0, 0, 0, 0, 0, 3), c(0, 0, 0, 0, 0, 0, 3))
para = cbind(c(12, 0.8, 0.8, 0.8, 0.8, 0.8, 0.8), c(12, 0.8, 0.8, 0.8, 0.8, 0.8, 0.8))
mu = c(0.2, 0.2)
s = 7
season_lens = 6*c(2, 4, 8, 16, 20)+1
n_seasons = 50

# Low intraspecific competition, symmetric

vlowintra = invasion_multi(n_seasons, init_pop, para, c(B, 0.10, 0.10), delta_s_lst, var = 0, season_lens)

# Mid intraspecific competition, asymmetric

vmidintra = invasion_multi(n_seasons, init_pop, para, c(B, 0.10, 0.12), delta_s_lst, var = 0, season_lens)

# High intraspecific competition, symmetric

vhighintra = invasion_multi(n_seasons, init_pop, para, c(B, 0.12, 0.12), delta_s_lst, var = 0, season_lens)

```

- Visualization

```{r}

vpllow = 
vlowintra %>% select(inv_i, lambda_ij, delta_s, season_len) %>% 
  pivot_wider(names_from = inv_i, values_from = lambda_ij) %>% 
  mutate(outcome = ifelse(`1`>0 & `2`>0, 3, 
                          ifelse(`1`>0 & `2`<0, 1, 
                                 ifelse(`1`<0 & `2`>0, 2, 4)))) %>% 
  ggplot(aes(x = as.factor(delta_s), y = as.factor(season_len), fill = factor(outcome))) + geom_tile(color = "white") + theme_bw() + 
  scale_x_discrete(name = "Initial stage difference", breaks = delta_s_lst, label = delta_s_lst) +
  scale_y_discrete(name = "# Generations per season", breaks = season_lens, label = (season_lens-1)/6) +
  # change here 
  # scale_fill_manual(name = "Outcome", labels = c("1 wins", "2 wins", "Coexistence", "Numeric PE"), values = color_scheme, guide = "none") +
  scale_fill_manual(name = "Outcome", labels = c("1 wins", "2 wins", "Numeric PE"), values = c("#a61b29", "#0eb0c9", "#577C8A"), guide = "none") +
  theme(panel.grid.major = element_blank()) + ggtitle("A") +
  theme(axis.text = element_text(size = 15), axis.ticks = element_blank(), axis.title = element_blank(),  
        legend.text = element_text(size = 17), legend.title = element_text(size = 20), plot.title = element_text(size = 20))

vplmid = 
vmidintra %>% select(inv_i, lambda_ij, delta_s, season_len) %>% 
  pivot_wider(names_from = inv_i, values_from = lambda_ij) %>% 
  mutate(outcome = ifelse(`1`>0 & `2`>0, 3, 
                          ifelse(`1`>0 & `2`<0, 1, 
                                 ifelse(`1`<0 & `2`>0, 2, 4)))) %>% 
  ggplot(aes(x = as.factor(delta_s), y = as.factor(season_len), fill = factor(outcome))) + geom_tile(color = "white") + theme_bw() + 
  scale_x_discrete(name = "Initial stage difference", breaks = delta_s_lst, label = delta_s_lst) +
  scale_y_discrete(name = "# Generations per season", breaks = season_lens, label = (season_lens-1)/6) +
  # change here 
  # scale_fill_manual(name = "Outcome", labels = c("1 wins", "2 wins", "Coexistence", "Numeric PE"), values = color_scheme, guide = "none") +
  scale_fill_manual(name = "Outcome", labels = c("1 wins", "2 wins", "Numeric PE"), values = c("#a61b29", "#0eb0c9", "#577C8A")) +
  theme(panel.grid.major = element_blank()) + ggtitle("B") +
  theme(axis.text = element_text(size = 15), axis.ticks = element_blank(), 
        axis.title.x = element_blank(), axis.title.y = element_text(size = 20), 
        legend.text = element_text(size = 17), legend.title = element_text(size = 20), plot.title = element_text(size = 20))

vplhigh = 
vhighintra %>% select(inv_i, lambda_ij, delta_s, season_len) %>% 
  pivot_wider(names_from = inv_i, values_from = lambda_ij) %>% 
  mutate(outcome = ifelse(`1`>0 & `2`>0, 3, 
                          ifelse(`1`>0 & `2`<0, 1, 
                                 ifelse(`1`<0 & `2`>0, 2, 4)))) %>% 
  ggplot(aes(x = as.factor(delta_s), y = as.factor(season_len), fill = factor(outcome))) + geom_tile(color = "white") + theme_bw() + 
  scale_x_discrete(name = "Initial stage difference", breaks = delta_s_lst, label = delta_s_lst) +
  scale_y_discrete(name = "# Generations per season", breaks = season_lens, label = (season_lens-1)/6) +
  # change here 
  # scale_fill_manual(name = "Outcome", labels = c("1 wins", "2 wins", "Coexistence", "Numeric PE"), values = color_scheme, guide = "none") +
  scale_fill_manual(name = "Outcome", labels = c("1 wins", "2 wins", "Numeric PE"), values = c("#a61b29", "#0eb0c9", "#577C8A"), guide = "none") +
  theme(panel.grid.major = element_blank()) + ggtitle("C") +
  theme(axis.text = element_text(size = 15), axis.ticks = element_blank(), 
        axis.title.y = element_blank(), axis.title.x = element_text(size = 20), 
        legend.text = element_text(size = 17), legend.title = element_text(size = 20), plot.title = element_text(size = 20))

vpllow / vplmid / vplhigh + plot_layout(guides = "collect")

```

## Figure S3

```{r}

ggsave("Figures/FigS3.png", plot = vpllow / vplmid / vplhigh + plot_layout(guides = "collect"), 
       width = 2250, height = 3375, unit = "px")

```

# Appendix IV

```{r}

B = 0.225
scal = 0.85
xmid = 0

adult_all = data.frame()
for (ds in -(1:4)) {
  for (season_len in c(13, 49)) {
    adult = convert_output(seasons(20*season_len, init_pop, para, alpha, ds, var = 0, season_len)) %>% 
            mutate(season_no = (time %/% season_len) + 1) %>% 
            group_by(species, season_no) %>% summarize(adult = sum(`6`)) %>%
            mutate(delta_s = ds, season_len = season_len)
    adult_all = rbind(adult_all, adult)
  }
}

plavgadp =
adult_all %>% filter(species == 2) %>% 
  ggplot(aes(x = season_no, y = log(adult+1), color = factor(delta_s))) + geom_line() + 
  geom_point(size = 3, alpha = 0.5) + 
  scale_color_manual(values = shades_kql, name = expression(paste(Delta, "s"))) + 
  scale_shape_manual(name = "T", values = c(16, 17)) + 
  scale_linetype_manual(name = "T", values = c(1, 8)) + ylab("log(adult population+1)") + 
  scale_x_continuous(name = "Number of seasons", breaks = c(1, 5, 10, 15), limits = c(1, 15)) +
  facet_grid( ~ season_len, scales = "free_y", 
             labeller = labeller(season_len = c(`13`="Two generations", `49`="Eight generations"))) +
  theme_bw() +
  theme(panel.grid.minor = element_blank(),
        axis.text = element_text(size = 15), axis.title = element_text(size = 20), 
        axis.ticks.y = element_blank(), plot.title = element_text(size = 20), 
        legend.text = element_text(size = 17), legend.title = element_text(size = 20), 
        strip.text = element_text(size = 15))

```

## Figure S4

```{r}

ggsave("Figures/FigS4.png", plot = plavgadp, 
       width = 2000, height = 1200, unit = "px")

```
